---
title: "ROH_summary_statistics"
output: html_document
date: "2025-06-11"
---

### This file contains code for calculating ROH summary statistics for worldwide and Near Oceanic populations
```{r}
options(scipen = 999)
library(ggplot2)
library(gridExtra)
library(RColorBrewer)
library(ggsci)
library(data.table)
library(tidyverse)
library(purrr)
library(cowplot) 
library(ggplotify)
library(ggtext)
library(dunn.test)
library(stats)
```

### Create .hom and .hom.indiv files
```{bash}
#!/bin/bash
#SBATCH --job-name={prefix}_MAF0.01_LDpruned_ROH
#SBATCH --mem=4G
#SBATCH --cpus-per-task=10
#SBATCH --time=30:00
#SBATCH --output={prefix}_MAF0.01_LDpruned_ROH_500Kb.log

#replace with specific PLINK version (PLINK 1.9)
PLINK="PLINK/1.9b_6.21-x86_64"

#prefix for LD and MAF0.01 pruned bed/bim/fam files 
#contaminated/related/etc. individuals were removed prior to pruning
prefix="./[input_file]"

out_prefix="global_autosomes"

module load ${PLINK}

plink --threads 10 \
--not-chr 23 \
--homozyg \
--homozyg-kb 500 \
--homozyg-snp 50 \
--homozyg-window-snp 50 \
--homozyg-window-het 0 \
--homozyg-window-missing 5 \
--homozyg-window-threshold 0.05 \
--homozyg-density 5000 \
--homozyg-gap 5000 \
--bfile ${prefix} \
--out ${out_prefix}_500Kb

```


```{r}
#read in .hom and .hom.indiv files
hom_indiv <- read.table("./data_files/[filename].hom.indiv", header = TRUE)

hom <- read.table("./data_files/[filename].hom", header = TRUE)

#metadata file containing sampleIDs, Region, Population information
#contaminated/related/etc. individuals have been excluded from this metadata file
#Vanuatu was already removed in the MAF and LD pruning step in PLINK, but this is just an extra step to make sure
metadata <- read.table("./[filename].tsv", sep = "\t", header = TRUE, fill = TRUE) %>%
  filter(Island != "Vanuatu" ) %>%
  dplyr::select(SampleID, Region, Population, Island)

#Oceania IDs only
OCN_included_samps <- metadata %>% filter(Region %in% "Oceania") 

#merge with hom.indiv
global_indiv <- merge(hom_indiv, metadata, by.x = "IID", by.y = "SampleID") %>%
  dplyr::select(IID, NSEG, KB, Region, Population, Island) %>%
  mutate(MB = KB*0.001)

#centromere regions
centromere_bed <- read.table("./data_files/hs37d5_hg19_centromeres.bed") %>%
  `colnames<-`(c("CHR", "centro_POS1", "centro_POS2"))

#remove from hom file
no_centro_hom <- centromere_bed %>%
  filter(CHR != c("X", "Y")) %>%
  mutate(CHR = as.numeric(CHR), centro_POS1 = as.numeric(centro_POS1), centro_POS2 = as.numeric(centro_POS2)) %>%
  left_join(hom, ., by = "CHR") %>%
  filter(POS1 > centro_POS2 | POS2 < centro_POS1+1) #add one to convert from BED coords

#add in metadata info
global_no_centro_hom <- merge(no_centro_hom, metadata, by.x = "IID", by.y = "SampleID") %>% 
  dplyr::select(-centro_POS1, -centro_POS2) %>%
  mutate(MB = KB*0.001)

#merged Oceanic groupings
renbell <- c("Rennell|Bellona")
lavonmus <- c("Lavongai|Mussau")
nainot <- c("Nailik|Notsi|Tigak")
nakman <- c("Nakanai|Mangseng")
sepgor <- c("Sepik|Goroka")

#replace names in df with groupings
global_no_centro_hom_merge <- global_no_centro_hom %>%
  mutate(Population = str_replace_all(Population, lavonmus, "Lavongai-Mussau"), Island = str_replace_all(Island, 'Mussau', "New Hanover")) %>%
  mutate(Population = str_replace_all(Population, nainot, "Nailik-Notsi-Tigak")) %>%
  mutate(Population = str_replace_all(Population, renbell, "Rennell-Bellona")) %>%
  mutate(Population = str_replace_all(Population, nakman, "Nakanai-Mangseng")) %>%
  mutate(Population = str_replace_all(Population, sepgor, "Goroka-Sepik")) %>%
  mutate(Population = str_replace_all(Population, c('Santa Cruz' = "Santa-Cruz", 'Vella Lavella' = "Vella-Lavella")))


# write to table 
#write.table(global_no_centro_hom_merge, "./data_files/global_MAF_LD_no_centro_hom_500Kb.tsv", sep = "\t", col.names = TRUE)

#summarize sROH and nROH per individual (creates a df that is pretty much the same as the hom.indiv file)
global_no_centro_hom_summary <- global_no_centro_hom_merge %>%
  group_by(IID, Population, Region, .drop = FALSE) %>%
  summarise(sum_MB = sum(MB), nseg = n())
```

### visualize sROH
```{r}
global_region_levels <- c("Africa", "Middle_East", "Europe", "Central_South_Asia", "East_Asia", "America", "Island_Southeast_Asia", "Oceania")

global_pop_levels <- c("Mbuti", "Biaka", "Mandenka", "Yoruba", "San", "BantuSouthAfrica", "BantuKenya",
                       "Druze", "Bedouin", "Palestinian", "Mozabite",
                       "French", "Sardinian", "Orcadian", "Russian", "BergamoItalian", "Tuscan", "Basque", "Adygei",
                       "Brahui", "Balochi", "Hazara", "Makrani", "Sindhi", "Pathan", "Kalash", "Burusho", "Uygur",
                       "Atayal", "Paiwan", "Cambodian", "Japanese", "Han", "Yakut", "Tujia", "Yi", "Miao", "Oroqen", 
                       "Daur", "Mongolian", "Hezhen", "Xibo", "NorthernHan", "Dai", "Lahu", "She", "Naxi", "Tu",
                       "Colombian", "Surui", "Maya", "Karitiana", "Pima", 
                       "Agta", "Cebuano", "Rampasasa",
                       "Goroka-Sepik", "Kove", "Nakanai-Mangseng","Mamusi","Ata","Melamela", "Baining-Kagat","Baining-Mali",
                       "Lavongai-Mussau", "Nailik-Notsi-Tigak", "Nasioi", "Saposa", "Vella-Lavella", "Malaita", "Santa-Cruz",
                       "Rennell-Bellona", "Tikopia")

## function to get the names of the segments in the plot with correct spacing
addline_format <- function(x,...){
    gsub('//s','/n',x)
}

global_no_centro_hom_summary$Region <- factor(global_no_centro_hom_summary$Region, levels = global_region_levels)
global_no_centro_hom_summary$Population <- factor(global_no_centro_hom_summary$Population, levels = global_pop_levels)

ggplot(data = global_no_centro_hom_summary) +
  geom_boxplot(mapping = aes(x = Population, y = sum_MB, fill = Region),
               alpha = 0.9,
               size = 0.25,
               outlier.size = 0.6) +
  annotate("segment", x=0, y=-130, xend=7.5, 
           yend=-130, size=1.5, col="#FF8966") +
  annotate("text", x = 3.75, y = -150, label = "Africa", size = 13/.pt, angle = 45) +
  annotate("segment", x=7.6, y=-130, xend=11.5, 
           yend=-130, size=1.5, col = "#CA2E55") +
  annotate("text", x = 9, y = -160, label = addline_format("Middle East"), 
           size = 13/.pt, angle = 45) +
  annotate("segment", x=11.6, y=-130, xend=18.9, 
           yend=-130, size=1.5, col = "#F6AE2D") +
  annotate("text", x = 14, y = -150, label = "Europe", 
           size = 13/.pt, angle = 45) +
  annotate("segment", x=19, y=-130, xend=28.5, 
           yend=-130, size=1.5, col = "#EB7BC0") +
  annotate("text", x = 23.5, y = -170, label = addline_format("Central/ South Asia"), 
           size = 13/.pt, angle = 45) +
  annotate("segment", x=28.6, y=-130, xend=48.9, 
           yend=-130, size=1.5, col = "#60992D") +
  annotate("text", x = 37.5, y = -155, label = "East Asia", 
           size = 13/.pt, angle = 45) +
  annotate("segment", x=49, y=-130, xend=53.9, 
           yend=-130, size=1.5, col = "#5887FF") +
  annotate("text", x = 50, y = -150, label = "America", 
           size = 13/.pt, angle = 45) +
  annotate("segment", x=54, y=-130, xend=56.9, 
           yend=-130, size=1.5, col = "#7E1F86") +
  annotate("text", x = 55, y = -157, label = addline_format("Island SEA"), 
           size = 13/.pt, angle = 45) +
  annotate("segment", x=57, y=-130, xend=73, 
           yend=-130, size=1.5, col = "#95D9DA") +
  annotate("text", x = 65.5, y = -150, label = "Oceania", size = 13/.pt, angle = 45) +
  scale_fill_manual(values = c("#FF8966", "#CA2E55","#F6AE2D","#EB7BC0","#60992D",
                                         "#5887FF","#7E1F86","#95D9DA")) +
  labs(y = expression("Cumulative Sum ROH (Mbp)")) +
  theme_bw() +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 9),
        axis.text.y = element_text(size = 9),
        axis.title.x=element_blank(),
        axis.title.y=element_text(size = 11),
        panel.grid.major = element_blank(),
        plot.margin = unit(c(0.5,1,5,1), "lines"),
        panel.border = element_rect(linewidth = 1)) +
  coord_cartesian(clip = "off", ylim = c(0, 700))

#ggsave("./figures/Fig1_panel_global_all_ROH.pdf", plot = last_plot(), width = 16, height = 10, units = "cm")
#ggsave("./figures/Fig1_panel_global_all_ROH.png", plot = last_plot(), width = 16, height = 10, units = "cm")
```


### Quantitative measurements of sROH and nROH globally
```{r}

#summarise total, means, and medians for sROH and nROH per population
pop_summary <- global_no_centro_hom_summary %>%
  group_by(Population, Region, .drop = FALSE) %>%
  summarise(pop_sum_MB = sum(sum_MB), 
            pop_nseg = sum(nseg),
            avg_sum_MB = mean(sum_MB),
            avg_nseg = mean(nseg),
            median_sum_MB = median(sum_MB),
            median_nseg = median(nseg),
            min_sum_MB = min(sum_MB),
            max_sum_MB = max(sum_MB),
            min_nseg = min(nseg),
            max_nseg = max(nseg),
            n = n_distinct(IID)) %>%
  left_join(., global_fROH_avg, by = c("Population", "Region")) %>%
  mutate(Region = factor(Region, levels = global_region_levels)) %>%
  arrange()

#summarise per region 
region_summary <- global_no_centro_hom_summary %>%
  group_by(Region, .drop = FALSE) %>%
  summarise(pop_sum_MB = sum(sum_MB), 
            pop_nseg = sum(nseg),
            avg_sum_MB = mean(sum_MB),
            avg_nseg = mean(nseg),
            median_sum_MB = median(sum_MB),
            median_nseg = median(nseg),
            min_sum_MB = min(sum_MB),
            max_sum_MB = max(sum_MB),
            min_nseg = min(nseg),
            max_nseg = max(nseg),
            n = n_distinct(IID)) %>%
  mutate(avg_MB = pop_sum_MB/n, 
         avg_nseg = pop_nseg/n, 
         Region = factor(Region, levels = global_region_levels)) %>%
  arrange()

#calculate variance per region
region_var <- global_no_centro_hom_summary %>%
  group_by(Region, .drop = FALSE) %>%
  summarise(sd_sROH = sd(sum_MB),
            sd_nROH = sd(nseg),
            var_sROH = var(sum_MB),
            var_nROH = var(nseg)) %>%
  arrange(var_sROH)

pop_summary_format <- pop_summary %>%
  dplyr::select(-pop_sum_MB, -pop_nseg) %>%
  mutate(avg_sum_MB = round(avg_sum_MB, 3),
         avg_nseg = round(avg_nseg, 3),
         min_sum_MB = round(min_sum_MB, 3),
         max_sum_MB = round(max_sum_MB, 3),
         median_sum_MB = round(median_sum_MB, 3),
         avg_fROH = round(avg_fROH, 3),
         min_fROH = round(min_fROH, 3),
         max_fROH = round(max_fROH, 3),
         median_fROH = round(median_fROH, 3)) %>%
  mutate(minmax_sROH = paste0(min_sum_MB, "-", max_sum_MB),
         minmax_nROH = paste0(min_nseg, "-", max_nseg)) %>%
  mutate(med_range_sROH = paste0(median_sum_MB, " ", "(", minmax_sROH, ")"),
         med_range_nROH = paste0(median_nseg, " ", "(", minmax_nROH, ")"),
         Population = factor(Population, levels = global_pop_levels),
         Region = factor(Region, levels = global_region_levels)) %>%
  arrange(Region) %>%
  dplyr::select(Population, Region, n, avg_sum_MB, 
                avg_nseg, med_range_sROH, med_range_nROH, median_fROH, max_fROH) %>%
  relocate(n, .after = Region) %>%
  `colnames<-`(c('Population',
                 'Region',
                 'n',
                 'Mean sROH',
                 'Mean nROH',
                 'Median sROH (Range)',
                 'Median nROH (Range)',
                 'Median FROH',
                 'Max fROH'
                 ))

#write.table(pop_summary_format, "./figures/TableS2_global_sum_stats.tsv", sep = "\t", quote = FALSE, row.names = FALSE, col.names = TRUE)
#length of the unmasked autosomal genome

```

### Global FROH sum stats
```{r}
total_auto <- 2684.673005

# filter above 1.5Mb according to McQuillan et al and divide by total autosomal length
# 1170 total individuals, so some had no calls over 1.5Mb
global_fROH <- global_no_centro_hom_merge %>%
  filter(MB >= 1.5) %>%
  group_by(IID, Population, Region, .drop = FALSE) %>%
  summarise(sum_MB = sum(MB), nseg = n()) %>%
  mutate(FROH = sum_MB/total_auto) %>%
  mutate(FROH_percent = FROH *100) %>%
  ungroup()

global_fROH$Region <- factor(global_fROH$Region, levels = global_region_levels)
global_fROH$Population <- factor(global_fROH$Population, levels = global_pop_levels)

#fROH measurements
fROH_pop_summary <- global_fROH %>%
  group_by(Population, Region) %>%
  summarise(avg_fROH = mean(FROH),
            min_fROH = min(FROH),
            max_fROH = max(FROH),
            median_fROH = median(FROH))

#avg fROH per region
fROH_region_summary <- global_fROH %>%
  group_by(Region) %>%
  summarise(avg_fROH = mean(FROH),
            min_fROH = min(FROH),
            max_fROH = max(FROH),
            median_fROH = median(FROH))

```

### Sum stats for Oceania only
```{r}
Oceania_pop_summary <- global_no_centro_hom_summary %>%
  filter(Region == "Oceania") %>%
  dplyr::select(-pop_sum_MB, -pop_nseg) %>%
  mutate(avg_sum_MB = round(avg_sum_MB, 3),
         avg_nseg = round(avg_nseg, 3),
         min_sum_MB = round(min_sum_MB, 3),
         max_sum_MB = round(max_sum_MB, 3),
         median_sum_MB = round(median_sum_MB, 3),
         avg_fROH = round(avg_fROH, 3),
         min_fROH = round(min_fROH, 3),
         max_fROH = round(max_fROH, 3),
         median_fROH = round(median_fROH, 3)) %>%
  mutate(minmax_sROH = paste0(min_sum_MB, "-", max_sum_MB),
         minmax_nROH = paste0(min_nseg, "-", max_nseg)) %>%
  mutate(med_range_sROH = paste0(median_sum_MB, " ", "(", minmax_sROH, ")"),
         med_range_nROH = paste0(median_nseg, " ", "(", minmax_nROH, ")"),
         Population = factor(Population, levels = global_pop_levels),
         Region = factor(Region, levels = global_region_levels)) %>%
  arrange(Region) %>%
  dplyr::select(Population, Region, n, avg_sum_MB, 
                avg_nseg, med_range_sROH, med_range_nROH, median_fROH, max_fROH) %>%
  relocate(n, .after = Region) %>%
  `colnames<-`(c('Population',
                 'Region',
                 'n',
                 'Mean sROH',
                 'Mean nROH',
                 'Median sROH (Range)',
                 'Median nROH (Range)',
                 'Median FROH',
                 'Max fROH'
                 ))

#write.table(Oceania_pop_summary, "./figures/Table1_OCN_sum_stats.tsv", 
#            sep = "\t", quote = FALSE, row.names = FALSE, col.names = TRUE)

```

### Summarize per length category, global
```{r}

#global populations, separate into length groups
global_length_groups <- global_no_centro_hom %>%
  group_by(Population) %>%
  mutate(group = case_when(
    MB <= 1 ~ "0.5-1",
    MB > 1 & MB <= 2 ~ "1-2", 
    MB > 2 & MB <= 4 ~ "2-4",
    MB > 4 & MB <= 8 ~ "4-8", 
    MB > 8 & MB <= 16 ~ "8-16",
    MB > 16 & MB <= 32 ~ "16-32",
    MB > 32 ~ "32+"
  ))

global_length_groups$group <- factor(global_length_groups$group, 
                                      levels = level_order <- c("0.5-1","1-2", "2-4", 
                                                                "4-8", "8-16", "16-32", "32+"))

global_length_groups$Population <- factor(global_length_groups$Population, levels = global_pop_levels)
global_length_groups$Region <- factor(global_length_groups$Region, levels = global_region_levels)


#summarize sum stats
global_length_group_summary <- global_length_groups %>%
  group_by(IID, group, .drop = FALSE) %>% #.drop = FALSE to keep the 0 values
  summarise(totalMB_per_cat = sum(MB), 
            total_nseg_per_cat = n()) %>%
  left_join(., metadata, by = c("IID" = "SampleID")) %>%
  dplyr::select(IID, group, totalMB_per_cat, total_nseg_per_cat, Region, Population) %>%
  group_by(group, Region) %>%
  summarise(mean_sROH = mean(totalMB_per_cat),
            min_sROH = min(totalMB_per_cat),
            max_sROH = max(totalMB_per_cat),
            sd_sROH = sd(totalMB_per_cat),
            mean_nROH = mean(total_nseg_per_cat),
            min_nROH = min(total_nseg_per_cat),
            max_nROH = max(total_nseg_per_cat),
            sd_nROH = sd(total_nseg_per_cat))

```


#summarize per length catergory, Oceania specific
```{r}
#Oceania pop levels
island_levels <- c("Goroka-Sepik", "Kove", "Nakanai-Mangseng","Mamusi","Ata","Melamela", "Baining-Kagat","Baining-Mali",
                   "Lavongai-Mussau", "Nailik-Notsi-Tigak", "Nasioi", "Saposa", "Vella-Lavella", "Malaita", "Santa-Cruz", "Rennell-Bellona", "Tikopia")

Oceania_length_groups <- global_no_centro_hom %>%
  filter(Region == "Oceania") %>%
  group_by(Population) %>%
  mutate(group = case_when(
    MB <= 1 ~ "0.5-1",
    MB > 1 & MB <= 2 ~ "1-2", 
    MB > 2 & MB <= 4 ~ "2-4",
    MB > 4 & MB <= 8 ~ "4-8", 
    MB > 8 & MB <= 16 ~ "8-16",
    MB > 16 & MB <= 32 ~ "16-32",
    MB > 32 ~ "32+"
  ))

Oceania_length_groups$group <- factor(Oceania_length_groups$group, levels = level_order <- c("0.5-1","1-2", "2-4", 
                                                                                             "4-8", "8-16", "16-32", "32+"))
Oceania_length_groups$Population <- factor(Oceania_length_groups$Population, levels = island_levels)

Oceania_length_group_summary <- Oceania_length_groups %>%
  group_by(IID, group, .drop = FALSE) %>% #.drop = FALSE to keep the 0 values
  summarise(totalMB_per_cat = sum(MB), 
            total_nseg_per_cat = n()) %>%
  left_join(., included_samps, by = c("IID" = "SampleID")) %>%
  mutate(Population = str_replace_all(Population, lavonmus, "Lavongai-Mussau"), Island = str_replace_all(Island, 'Mussau', "New Hanover")) %>%
  mutate(Population = str_replace_all(Population, nainot, "Nailik-Notsi-Tigak")) %>%
  mutate(Population = str_replace_all(Population, renbell, "Rennell-Bellona")) %>%
  mutate(Population = str_replace_all(Population, nakman, "Nakanai-Mangseng")) %>%
  mutate(Population = str_replace_all(Population, sepgor, "Goroka-Sepik")) %>%
  mutate(Population = str_replace_all(Population, c('Santa Cruz' = "Santa-Cruz", 'Vella Lavella' = "Vella-Lavella"))) %>%
  dplyr::select(IID, group, totalMB_per_cat, total_nseg_per_cat, Region, Population) %>%
  group_by(Population, group, .drop = FALSE) %>%
  summarise(mean_sROH = mean(totalMB_per_cat),
            median_sROH = median(totalMB_per_cat),
            sd_sROH = sd(totalMB_per_cat),
            mean_nROH = mean(total_nseg_per_cat),
            median_nROH = median(total_nseg_per_cat),
            sd_nROH = sd(total_nseg_per_cat))

Oceania_length_group_summary$Population <- factor(Oceania_length_group_summary$Population, levels = island_levels)


#how many individuals have ROH above a certain length category
OCN_ind_4to8 <- Oceania_length_groups %>%    
  filter(group == "4-8") %>%
  summarise(ind = unique(IID))

OCN_ind_8to16 <- Oceania_length_groups %>%   
  filter(group == "8-16") %>%
  summarise(ind = unique(IID))

OCN_ind_16to32 <- Oceania_length_groups %>%  
  filter(group == "16-32") %>%
  summarise(ind = unique(IID))

#encode shapes for plotting
island_shapes <- c(17,16,18,17,15,18,15,16,15,16,15,16,15,17,16,16,18)

#color scheme
# 8 total island groups, this is broken up more so by ancestry components + geography
island_col2 <- c("#225ea8", "#AF96FF", "#AF96FF", "#c7e9b4", "#c7e9b4", "#c7e9b4", "#7fcd88", "#7fcd88", "#41b6c4", "#41b6c4","#FBE83A", "#FBE83A", "#90CBEF", "#90CBEF", "#90CBEF", "#544287", "#544287")

OCN_sROH_lineplot <- Oceania_length_group_summary %>%
  ggplot() +
  geom_line(aes(x = group, y = median_sROH, col = Population, group = Population), size = 1.2) +
  geom_point(aes(x = group, y = median_sROH, col = Population, fill = Population, shape = Population), size = 4) +
  scale_color_manual(values = island_col) +
  scale_fill_manual(values = island_col) +
  scale_shape_manual(values = island_shapes) + 
  labs(y = expression("Cumulative sROH"), x = "Length Category (Mbp)") +
  scale_x_discrete(expand = c(0,0)) +
  theme_bw() +
  theme(axis.title.x = element_text(size = 11),
        legend.title = element_blank(),
        legend.text = element_text(size = 10),
        axis.title.y = element_text(size = 11),
        axis.text.x = element_text(size = 11),
        axis.text.y = element_text(size = 11),
        legend.position = "none",
        panel.border = element_rect(linewidth = 1))
OCN_sROH_lineplot

#ggsave("../figures/Figure2_OCN_sROH_lineplot_allcol.pdf", OCN_sROH_lineplot, width = 12, height = 10, unit = "cm")
#ggsave("../figures/Figure2_OCN_sROH_lineplot_allcol.png", OCN_sROH_lineplot, width = 12, height = 10, unit = "cm")
```

### Kruskal-Wallis-Dunn test to compare medians worldwide
```{r}
#by region
global_region_KW <- global_no_centro_hom %>%
  group_by(IID, Region, .drop = FALSE) %>%
  summarise(sum_MB = sum(MB), nseg = n()) %>%
  mutate(Region = factor(Region, levels = global_region_levels)) %>%
  arrange(Region)

#KW 
kruskal.test(sum_MB ~ Region, data = global_region_KW)
kruskal.test(nseg ~ Region, data = global_region_KW)

#dunn test by region
dunn_total_sumMB <- dunn.test(x = global_region_KW$sum_MB,    #sROH
          g = global_region_KW$Region,
          method = "bonferroni",
          kw = TRUE,
          label = TRUE,
          table = TRUE,
          wrap = TRUE)
dunn_total_nseg <- dunn.test(x = global_region_KW$nseg,    #nROH
          g = global_region_KW$Region,
          method = "bonferroni",
          kw = TRUE,
          label = TRUE,
          table = TRUE,
          wrap = TRUE)


#for plotting 
# by region
total_sumMB_df <- data.frame(compare1 = dunn_total_sumMB$comparisons,
                                   adj_pval = dunn_total_sumMB$P.adjusted) %>%
  separate(., col = compare1, c("pop1", "pop2"), sep = " - ") %>%
  mutate(neg_log10 = -log10(adj_pval))

total_sumMB_df <- bind_rows(total_sumMB_df, 
                                 total_sumMB_df %>%
                                   mutate(tmp1 = pop2, tmp2 = pop1) %>%
                                   dplyr::select(-c(pop1,pop2)) %>%
                                   mutate(pop1 = tmp1, pop2 = tmp2) %>%
                                   dplyr::select(-c(tmp1, tmp2)))

total_sumMB_df$pop1 <- factor(total_sumMB_df$pop1, levels = global_region_levels)
total_sumMB_df$pop2 <- factor(total_sumMB_df$pop2, levels = global_region_levels)


total_nseg_df <- data.frame(compare1 = dunn_total_nseg$comparisons,
                                   adj_pval = dunn_total_nseg$P.adjusted) %>%
  separate(., col = compare1, c("pop1", "pop2"), sep = " - ") %>%
  mutate(neg_log10 = -log10(adj_pval))

total_nseg_df <- bind_rows(total_nseg_df, 
                                 total_nseg_df %>%
                                   mutate(tmp1 = pop2, tmp2 = pop1) %>%
                                   dplyr::select(-c(pop1,pop2)) %>%
                                   mutate(pop1 = tmp1, pop2 = tmp2) %>%
                                   dplyr::select(-c(tmp1, tmp2)))

total_nseg_df$pop1 <- factor(total_nseg_df$pop1, levels = global_region_levels)
total_nseg_df$pop2 <- factor(total_nseg_df$pop2, levels = global_region_levels)

#plots
total_sumMB_plot <- total_sumMB_df %>% ggplot() +
  geom_raster(aes(x = pop1, y = pop2, fill = neg_log10)) +
  scale_fill_gradient(low = "#BAD4C9", high = "#274739") + #low is sig pval, darker color
  scale_x_discrete(labels = c("Africa", "Middle_East", "Europe", "Central_South_Asia", "East_Asia", "America", "Island_Southeast_Asia", "Oceania")) +
  scale_y_discrete(labels = c("Africa", "Middle_East", "Europe", "Central_South_Asia", "East_Asia", "America", "Island_Southeast_Asia", "Oceania")) +
  theme_bw() +
  labs(fill = expression(paste("Adjusted","-log10",~"p-value"))) +
  theme(panel.background = element_blank(),
        panel.grid = element_blank(),
        axis.text.x = ggtext::element_markdown(angle = 90, vjust = .5, hjust = 0.9, size = 8),
        axis.text.y = ggtext::element_markdown(size = 8),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        legend.position = "none")

total_nseg_plot <- total_nseg_df %>% ggplot() +
  geom_raster(aes(x = pop1, y = pop2, fill = neg_log10)) +
  scale_fill_gradient(low = "#BAD4C9", high = "#274739") + #low is sig pval, darker color
  scale_x_discrete(labels = c("Africa", "Middle_East", "Europe", "Central_South_Asia", "East_Asia", "America", "Island_Southeast_Asia", "Oceania")) +
  scale_y_discrete(labels = c("Africa", "Middle_East", "Europe", "Central_South_Asia", "East_Asia", "America", "Island_Southeast_Asia", "Oceania")) +
  theme_bw() +
  labs(fill = expression(paste("Adjusted","-log10",~"p-value"))) +
  theme(panel.background = element_blank(),
        panel.grid = element_blank(),
        axis.text.x = ggtext::element_markdown(angle = 90, vjust = .5, hjust = 0.9, size = 8),
        axis.text.y = ggtext::element_markdown(size = 8),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        legend.position = "none")

#ggsave("./figures/all_dunn_global_region_plot.png", plot = dunn_global_region, width = 14, height = 10, units = "cm")
#ggsave("./figures/all_dunn_global_region_plot.pdf", plot = dunn_global_region, width = 14, height = 10, units = "cm")
```

### Kruskal-Wallis-Dunn test to compare medians in Oceania
```{r}
# gives sROH and nROH per individual in Oceania
Oceania_KW <- global_MAF_LD_no_centro_hom %>%
  filter(Region =="Oceania") %>%
  group_by(IID, Population, Region, .drop = FALSE) %>%
  summarise(sum_MB = sum(MB), nseg = n()) %>%
  mutate(Population = factor(Population, levels = island_levels))

#test for both sROH and nROH
#if p-val shows significant difference, move on to Dunn test
kruskal.test(sum_MB ~ Population, data = Oceania_KW)
kruskal.test(nseg ~ Population, data = Oceania_KW)

OCN_dunn_total_sumMB <- dunn.test(x = Oceania_KW$sum_MB,    #sROH
          g = Oceania_KW$Population,
          method = "bonferroni",
          kw = TRUE,
          label = TRUE,
          table = TRUE,
          wrap = TRUE)
OCN_dunn_total_NSEG <- dunn.test(x = Oceania_KW$nseg,    #nROH
          g = Oceania_KW$Population,
          method = "bonferroni",
          kw = TRUE,
          label = TRUE,
          table = TRUE,
          wrap = TRUE)

OCN_total_sumMB_df <- data.frame(adj_pval = OCN_dunn_total_sumMB$P.adjusted, compare1 = OCN_dunn_total_sumMB$comparisons)
OCN_total_sumMB_df <- OCN_total_sumMB_df %>%
  separate(., col = compare1, c("pop1", "pop2"), sep = " - ") %>%
  mutate(neg_log10 = -log10(adj_pval))

OCN_total_nseg_df <- data.frame(adj_pval = OCN_dunn_total_NSEG$P.adjusted, compare1 = OCN_dunn_total_NSEG$comparisons)
OCN_total_nseg_df <- OCN_total_nseg_df %>%
  separate(., col = compare1, c("pop1", "pop2"), sep = " - ") %>%
  mutate(neg_log10 = -log10(adj_pval))


#produce heatmap plots 

#sumMB for plotting 
OCN_total_sumMB_df <- data.frame(compare1 = OCN_dunn_total_sumMB$comparisons,
                                   adj_pval = OCN_dunn_total_sumMB$P.adjusted) %>%
  separate(., col = compare1, c("pop1", "pop2"), sep = " - ") %>%
  mutate(neg_log10 = -log10(adj_pval))

OCN_total_sumMB_df <- bind_rows(OCN_total_sumMB_df, 
                                 OCN_total_sumMB_df %>%
                                   mutate(tmp1 = pop2, tmp2 = pop1) %>%
                                   dplyr::select(-c(pop1,pop2)) %>%
                                   mutate(pop1 = tmp1, pop2 = tmp2) %>%
                                   dplyr::select(-c(tmp1, tmp2)))

OCN_total_sumMB_df$pop1 <- factor(OCN_total_sumMB_df$pop1, levels = island_levels)
OCN_total_sumMB_df$pop2 <- factor(OCN_total_sumMB_df$pop2, levels = island_levels)

#nseg for plotting
OCN_total_NSEG_df <- data.frame(compare1 = OCN_dunn_total_NSEG$comparisons,
                                   adj_pval = OCN_dunn_total_NSEG$P.adjusted) %>%
  separate(., col = compare1, c("pop1", "pop2"), sep = " - ") %>%
  mutate(neg_log10 = -log10(adj_pval))

OCN_total_NSEG_df<- bind_rows(OCN_total_NSEG_df, 
                                 OCN_total_NSEG_df %>%
                                   mutate(tmp1 = pop2, tmp2 = pop1) %>%
                                   dplyr::select(-c(pop1,pop2)) %>%
                                   mutate(pop1 = tmp1, pop2 = tmp2) %>%
                                   dplyr::select(-c(tmp1, tmp2)))

OCN_total_NSEG_df$pop1 <- factor(OCN_total_NSEG_df$pop1, levels = island_levels)
OCN_total_NSEG_df$pop2 <- factor(OCN_total_NSEG_df$pop2, levels = island_levels)


OCN_total_sumMB_plot <- OCN_total_sumMB_df %>% ggplot() +
  geom_raster(aes(x = pop1, y = pop2, fill = neg_log10)) +
  scale_fill_gradient(low = "#BAD4C9", high = "#274739") + #low is sig pval, darker color
  scale_x_discrete(labels = c("Goroka-Sepik", "Kove", "Nakanai-Mangseng","Mamusi","Ata","Melamela", "**Baining-Kagat**","**Baining-Mali**","Lavongai-Mussau", "Nailik-Notsi-Tigak", "Nasioi", "Saposa", "Vella-Lavella", "Malaita", "Santa-Cruz", "**Rennell-Bellona**", "Tikopia")) +
  scale_y_discrete(labels = c("Goroka-Sepik", "Kove", "Nakanai-Mangseng","Mamusi","Ata","Melamela", "**Baining-Kagat**","**Baining-Mali**","Lavongai-Mussau", "Nailik-Notsi-Tigak", "Nasioi", "Saposa", "Vella-Lavella", "Malaita", "Santa-Cruz", "**Rennell-Bellona**", "Tikopia")) +
  theme_bw() +
  labs(fill = expression(paste("Adjusted","-log10",~"p-value"))) +
  theme(panel.background = element_blank(),
        panel.grid = element_blank(),
        axis.text.x = ggtext::element_markdown(angle = 90, vjust = .5, hjust = 0.9, size = 8),
        axis.text.y = ggtext::element_markdown(size = 8),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        legend.position = "none")


OCN_total_NSEG_plot <- OCN_total_NSEG_df %>% ggplot() +
  geom_raster(aes(x = pop1, y = pop2, fill = neg_log10)) +
  scale_fill_gradient(low = "#BAD4C9", high = "#274739") + #low is sig pval, darker color
  scale_x_discrete(labels = c("Goroka-Sepik", "Kove", "Nakanai-Mangseng","Mamusi","Ata","Melamela", "**Baining-Kagat**","**Baining-Mali**","Lavongai-Mussau", "Nailik-Notsi-Tigak", "Nasioi", "Saposa", "Vella-Lavella", "Malaita", "Santa-Cruz", "**Rennell-Bellona**", "Tikopia")) +
  scale_y_discrete(labels = c("Goroka-Sepik", "Kove", "Nakanai-Mangseng","Mamusi","Ata","Melamela", "**Baining-Kagat**","**Baining-Mali**","Lavongai-Mussau", "Nailik-Notsi-Tigak", "Nasioi", "Saposa", "Vella-Lavella", "Malaita", "Santa-Cruz", "**Rennell-Bellona**", "Tikopia")) +
  labs(fill = expression(paste("Adjusted","-log10",~"p-value"))) +
  theme_bw() +
  theme(panel.background = element_blank(),
        panel.grid = element_blank(),
        axis.text.x = ggtext::element_markdown(angle = 90, vjust = .5, hjust = 0.9, size = 8),
        axis.text.y = ggtext::element_markdown(size = 8),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        legend.position = "none",
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 9))

#arrange into one plot
OCN_all_dunn_plot <- arrangeGrob(OCN_total_sumMB_plot, OCN_total_NSEG_plot, 
             ncol = 2, widths = c(0.5, 0.5))

#ggsave("./figures/OCN_all_dunn_plot.png", plot = OCN_all_dunn_plot, width = 14, height = 10, units = "cm")
#ggsave("./figures/OCN_all_dunn_plot.pdf", plot = OCN_all_dunn_plot, width = 14, height = 10, units = "cm")
```